Evaluation of the factor structure and structural validity of the Maltreatment and Abuse Chronology of Exposure (MACE) scale

### 2023-02-18 manuscript/main.tex
abbreviated & submitted

\documentclass[letterpaper,man,natbib,noextraspace,floatsintext,longtable,12pt]{apa6}

\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{booktabs}             % fancy latex tables
\usepackage[export]{adjustbox}    % center wide tables on page
\usepackage{setspace}             % adjust caption linespacing 
\usepackage{multibib}             % references in supplement
\usepackage{hyperref}             % manage hyperlink colors
\usepackage{hanging}

@@

% define title page
\title{Examination and replication of the factorial validity and subscale reliability of the Maltreatment and Abuse Chronology of Exposure (MACE) scale}
\shorttitle{Factorial validity of the MACE}
\author{Samuel Zorowitz$^{1*}$, Lauri Tuominen$^{2}$}
\affiliation{$^1$Princeton Neuroscience Institute, Princeton University, USA\\$^2$The Royal’s Institute of Mental Health Research, University of Ottawa, Canada}

@@

\noindent *Corresponding author: zorowitz@princeton.edu

@@

\noindent \textbf{Objective}: The Maltreatment and Abuse Chronology of Exposure (MACE) scale is a retrospective self-report measure of childhood maltreatment. Despite increasing use, the factor structure of the MACE has not been previously investigated. As such, the psycho- metric properties of the MACE total and subscale scores are indeterminate. The aims of the current study was to examine the factorial validity of the MACE and the reliability of its total and subscale scores.

@@

\noindent \textbf{Methods}: We analyzed the response data from two independent samples of participants (N=1051 \& N=582) using confirmatory factor analysis. We fit a series of models with increasing dimensionality to explore factorial validity. Furthermore, we used model-based indices to estimate the reliability of the MACE total and subscale scores, and to test for essential unidimensionality.

@@

\noindent \textbf{Results}: Confirmatory factor analysis indicated that a bifactor S-1 model with two specific factors (peer victimization, reverse-scoring) provided the best and most parsimonious fit to the data. Despite this, model-based indices suggested the MACE scale is essentially unidimensional. We found the MACE total score was a reliable and valid measure of overall childhood maltreatment. In contrast, the MACE subscale scores exhibited little reliable variance after controlling for general maltreatment.

@@

Childhood maltreatment is an unfortunately common experience for children around the world \citep{stoltenborgh2015prevalence}. It is a risk factor for multiple poor outcomes later in life including the development of mental illness \citep{kessler2010childhood}, cognitive impairment \citep{su2019does}, and poor physical health \citep{wegman2009meta}. Childhood maltreatment also predicts worse treatment outcomes for several psychiatric disorders \citep{nanni2012childhood, thomas2019childhood}. Thus, the reliable and valid assessment of childhood maltreatment is an important research goal in order to refine our understanding of the link between maltreatment and health outcomes; augment treatment strategies for victims of childhood maltreatment; and ultimately design better interventions to prevent maltreatment in the first place.

There already exist many measures of childhood abuse and maltreatment for researchers to choose from \citep{saini2019systematic}. One recently developed measure, the Maltreatment and Abuse Chronology of Exposure scale (MACE; \citealt{teicher2015maltreatment}), has a number of notable advantages. First, the MACE is a 52-item retrospective self-report scale composed of 10 subscales, each of which measures a unique type of childhood maltreatment. The content of six of these subscales overlap with other childhood maltreatment scales (e.g. physical abuse, neglect), whereas the remaining four are less commonly represented in other scales (e.g. peer victimization, witnessing domestic violence). Second, using item response theory, the items in the MACE were carefully selected to measure maltreatment experiences of increasing severity in order to finely differentiate people with differing exposure levels. Third, the MACE also asks for the onset and duration of maltreatment experiences, which is necessary for investigating how the timing of abuse shapes development. Finally, the MACE was originally developed to provide two measures: a total score (number of overall maltreatment experiences) and a multiplicity score (number of types of maltreatment experienced). These scores have excellent temporal stability, convergent validity with other childhood maltreatment scales, and good predictive validity for a multitude of psychiatric symptoms \citep{teicher2015maltreatment}. For these reasons, the MACE has been included as among the best measures of childhood maltreatment currently available \citep{saini2019systematic, georgieva2022systematic}.

In recent years, there have been calls for childhood maltreatment research to move away from the cumulative risk framework \citep{evans2013cumulative} --- which focuses only on the number of adverse experiences --- towards a dimensional approach, which emphasizes how distinct types of maltreatment uniquely shape developmental trajectories \citep{mclaughlin2016beyond, belsky2012beyond}. In response, some researchers have begun to use the MACE to make new indices of maltreatment in lieu of, or in addition to, the total and multiplicity scores. For example, some researchers have used all 10 MACE subscale scores to study the unique contributions of each type of maltreatment to the development of psychopathology \citep{schalinski2015type, gerke2018childhood, schalinski2019early}. Citing a prominent dimensional model of maltreatment \citep{mclaughlin2014childhood}, others have made novel threat and deprivation subscale scores made up of items in the MACE that measure abuse and neglect experiences, respectively \citep{schalinski2018defining, schalinski2019environmental, teicher2018differential}. However, the psychometric properties of these subscale scores have not been thoroughly investigated.

A critical challenge in studying the consequences of childhood maltreatment is the high co-occurrence of these adverse experiences. That is, children who have experienced any single type of maltreatment are likely to have experienced multiple other types \citep{dong2004interrelatedness, herrenkohl2009assessing, kessler2010childhood}. As such, isolating the unique contribution of one type of maltreatment to an outcome is difficult insofar as any one type of adversity is likely confounded with overall maltreatment. Concretely, this means that researchers intending to study the sequela of particular types of maltreatment using subscale scores need to verify that those scores actually reflect a construct unique from overall maltreatment, lest they risk drawing spurious conclusions.

Bifactor models, a type of hierarchical factor model, are well-suited for addressing this issue \citep{bornovalova2020appropriate}. They assume that the covariation in responses to a set of items can be accounted for by a \emph{general} factor, reflecting shared variance across all items, and one or more \emph{specific} factors, reflecting additional shared variance among subsets of items \citep{Reise2012-ql}. Crucially, bifactor models can be used to calculate model-based reliability statistics, including the omega hierarchical ($\omega_h$) and omega subscale ($\omega_s$) indices \citep{reise2013scoring}, which respectively measure the proportion of variance in total and subscale scores attributable to the general factor and specific factors. Applied to the MACE, these indices can help determine if subscale scores measuring particular types of maltreatment are interpretable as such, or if they instead principally reflect overall maltreatment. In addition, bifactor models permit the calculation of other useful indices like the explained common variance (ECV; \citealt{sijtsma2009use}), which measures what fraction of the total reliable variance in a scale can be attributed to a general factor. Applied to the MACE, the ECV can quantify the degree to which the MACE is essentially unidimensional (i.e. summarizable by a one-factor model).

The overarching aim of the present study was to investigate the factor structure of the MACE, which has received little previous study to date \citep{saini2019systematic}. To do so, we fit and evaluated a series of confirmatory item factor models to the responses from two independent samples of participants who completed the MACE. The primary goal of our analyses was to quantify the reliability of the MACE total and subscale scores, and to determine whether the latter would provide unique information above and beyond the total score. A secondary goal was to examine the dimensionality of the MACE; that is, to determine whether responses on the MACE are best-explained by a hierarchical structure with 10 factors (one per subscale) or by a more parsimonious factor structure. All data and analysis materials are publicly available at: \url{https://github.com/szorowi1/mace-irt}.

@@

The participants in the current study (N=1633 total) belonged to one of two samples. The first sample of participants (N=1051) are from the original development of the MACE \citep{teicher2015maltreatment}. The full recruitment details for this sample have been reported elsewhere \citep{teicher2015maltreatment}. Briefly, these participants were recruited from the greater Boston area between 2010 and 2013. Inclusion criteria included being medically healthy, unmedicated, and between 18–25 years of age. This sample will hereafter be referred to as the \textit{original} sample.

The second sample of participants (N=687) were recruited from the Prolific Academic platform (\url{https://www.prolific.co}) to participate in an online experiment in August -- October, 2021. Of these, N=582 were retained for analysis (see Exclusion Criteria below). Participants were eligible for participation if they were 18 years or older and currently resided in the United States or Canada. Participants received monetary compensation for their time (rate: \$12 USD/hr). This study was approved by the Institutional Review Board of the University of Ottawa (\#2021003), and all participants provided informed consent. This sample will hereafter be referred to as the \textit{replication} sample.

\begin{table}[t!]
    \centering
    \begin{tabular*}{\textwidth}{lccc}
    \toprule
    Variable & Original (N=1051) & Replication (N=582) & p-value \\
    \midrule
    \textbf{Gender}, N (\%)                & & & p < 0.001 \\
    \hspace{1em} Women                     & 670 (63.7\%) & 310 (53.3\%) & \\
    \hspace{1em} Men                       & 381 (36.3\%) & 258 (44.3\%) & \\
    \hspace{1em} Transgender or NB         &   0  (0.0\%) &   9  (1.5\%) & \\
    \hspace{1em} Rather not say            &   0  (0.0\%) &   5  (0.9\%) & \\
    \midrule
    \textbf{Age}, yrs                      & & & p < 0.001 \\
    \hspace{1em} Mean (range)              & 23.1 (18 -- 27) & 30.2 (18 -- 80) & \\
    \midrule
    \textbf{Ethnicity}, N (\%)             & & & p = 0.214 \\
    \hspace{1em} Not Hispanic or Latino    & 963 (91.6\%) & 518 (89.0\%) & \\
    \hspace{1em} Hispanic or Latino        &  84  (8.0\%) &  57  (9.8\%) & \\
    \hspace{1em} Rather not say            &   4  (0.4\%) &   7  (1.2\%) & \\
    \midrule
    \textbf{Race}, N (\%)                  & & & p = 0.677 \\
    \hspace{1em} White                     & 806 (76.7\%) & 441 (75.8\%) & \\ 
    \hspace{1em} Asian                     & 103  (9.8\%) &  65 (11.2\%) & \\
    \hspace{1em} Black or African American &  81  (7.7\%) &  31  (5.3\%) & \\
    \hspace{1em} Other                     &  61  (5.8\%) &  28  (4.8\%) & \\
    \hspace{1em} Rather not say            &   0  (0.0\%) &  17  (2.9\%) & \\
    \bottomrule
    \end{tabular*}
    \caption{\normalfont Demographics of the two samples included in this study. The original sample was recruited and reported by \cite{teicher2015maltreatment}. The replication sample was recruited by the current authors for the purposes of this study.}
    \label{tab:demographics}
\end{table}

The demographics of the two samples are summarized in Table \ref{tab:demographics}. The majority of participants in both samples identified as women, though the original sample was composed of proportionally more woman than the replication sample ($z$ = 4.142, $p$ < 0.001). Participants in the original sample were also younger on average ($t$ = -19.313, $p$ < 0.001). The majority of participants in both samples identified as White and not as Hispanic or Latino. The proportion of participants identifying as White was not significantly different across samples ($z = 0.417$, $p = 0.677$), nor was the proportion of participants identifying as Hispanic or Latino ($z$ = -1.241, $p$ = 0.214).

@@

All participants completed the 52-item version of the MACE \citep{teicher2015maltreatment}. The MACE is a retrospective self-report scale that measures 10 unique types of childhood maltreatment (i.e. parental verbal abuse, physical abuse, nonverbal emotional abuse, sexual abuse, emotional neglect, physical neglect, witnessing violence to siblings, witnessing interparental violence, peer verbal abuse, peer physical abuse). For each item, participants endorse whether they ever experienced a particular event during childhood (e.g. ``Parents or guardians intentionally pushed, grabbed, shoved, slapped, pinched, punched or kicked you'') and, if so, at what ages (up to age 18). Here we only analyze the dichotomous endorsement responses (Yes = 1, No = 0) and not the chronology data. Six of the 52-items --- evenly divided between the emotional and physical neglect subscales --- are positively valenced and reverse-scored (e.g. ``One or more individuals in your family helped you feel important or special''). For these items we follow the scoring convention set by \cite{teicher2015maltreatment}: a score of 1 is assigned to a response only in the absence of a positive event for all 18 years of childhood. This is in contrast to all other items, where a score of 1 is assigned if maltreatment was experienced in at least one year of childhood.

In both the original and replication samples, participants completed a number of additional self-report symptom measures. We do not include these in our analyses and therefore do not discuss these measures further in the main text. For completeness, we report these secondary self-report measures in the supplementary materials.

@@

To ensure data quality in the replication sample, we relied on attention checks embedded in the self-report measures to identify and remove participants engaging in careless or insufficient effort responding (\citealt{zorowitz2021inattentive}; see the supplementary materials for the attention checks). The data from N=64 (9.9\%) participants in the replication sample were excluded for failing one or more of these attention checks, leaving the data from N=582 participants for analysis.

\subsection{Differential item functioning}

A mean shift in overall endorsement rates by sample is not necessarily evidence for a violation of measurement invariance. One possibility is that the replication sample had simply experienced more adversity than the original sample. A second is that the differences in the demographic composition of the two samples (e.g. gender, age) explain the differences in endorsement rates. To investigate these possibilities, we tested for uniform differential item functioning (DIF) using logistic regression, in which a participant's response on an item is regressed against their rest score (i.e. their total score for all other items), sample membership, gender, and age.

Using the cutoffs recommended by \cite{hidalgo2014binary}, we identified 11 of 52 items on the MACE with large DIF by study (Figure \ref{fig:dif_study}). These included all six reverse-scored items. We also found an additional 11 items with large DIF by gender (Figure \ref{fig:dif_gender}). Five of these items concerned sexual abuse (endorsed more frequently by women), and another five concerned peer physical abuse (endorsed more frequently by men). We identified no items that exhibited large DIF by age. Given the number of items with DIF by sample membership, we decided to analyze each sample separately.

\subsection{Item response dependence}

Structurally-dependent or redundant items can cause serious problems for item factor models. First, they can inflate the estimates of item discrimination parameters, making items appear more reliable than they actually are \citep{marais2008formalizing}. Second, response dependence can cause overestimation of item loadings on specific factors in bifactor models, making specific factors (and subscores) appear more reliable than they actually are \citep{reise2013applying}. Following \cite{marais2008formalizing}, we therefore combined each set of dependent items into a single polytomous item thereby reducing the MACE from 52 to 39 items in total.

@@

To investigate the factor structure of the MACE, we fit a series of four confirmatory item factor models separately to the response data from each sample. (The suitability of common factor models (CFMs) with reflective indicators for explaining childhood maltreatment data has been contested. For interested readers, we justify our use of CFMs to explain responses to the MACE in the supplement.) The basis of each model is the graded response model for ordered polytomous data \citep{samejima1997graded}. The factor structure for each model is depicted in Figure \ref{fig:models}. The first is a one-factor (undimensional) model where all items load onto a general maltreatment factor (Figure \ref{fig:models}a). The second is a bifactor model where all items load onto a general maltreatment factor and also one of 10 specific factors based on their subscale membership (Figure \ref{fig:models}b). This model will be used to evaluate if the MACE can produce reliable subscale scores for each of the 10 types of maltreatment assessed by the MACE.

The third model was motivated by a dimensional theory of childhood maltreatment that distinguishes between threat and deprivation experiences as two distinct types of adversity \citep{mclaughlin2014childhood}. To examine this model of the MACE, we initially fit a bifactor model where all items loaded onto a general maltreatment factor, and also a threat or neglect specific factor. However, we observed vanishing item loadings on the threat specific factor. Therefore, we instead fit a bifactor S-1 model \citep{eid2017anomalous} where all items load onto a general maltreatment factor, and only items measuring neglect (i.e. items from the physical and emotional neglect subscales) load onto a specific neglect factor (Figure \ref{fig:models}c). We fit this model to see if the MACE can be used to form neglect subscale scores that are reliable and distinguishable from general maltreatment.

@@

The fourth and final model we fit proposes a more parsimonious structure of the MACE based on the content of its items. Items on the MACE can be categorized according to the source of maltreatment: parents (or other adults in the home) and peers. Thus, we hypothesized there would be two factors underlying responses on the MACE: a primary parental maltreatment factor and a secondary peer victimization factor. From preliminary model fits, we also suspected the presence of a tertiary methods factor affecting the six reverse-scored items. We initially fit a bifactor model in which all items loaded onto a general maltreatment factor and one of three specific factors (parental maltreatment, peer victimization, reverse-scored); however, we observed vanishing loadings on the first specific factor. We instead fit a bifactor S-1 model where all items load onto a general maltreatment factor, and the peer victimization \& reverse-scored items additionally load onto their own specific factors (Figure \ref{fig:models}d). We fit this model in part to test a more parsimonious factor structure of the MACE, and to see if the MACE can be used to form peer victimization subscale scores that are reliable and distinguishable from general maltreatment.

@@

We relied on multiple indices to judge the fit of the models to the data. First we calculated three traditional goodness-of-fit statistics using the ordinal $M_2$ statistic \citep{cai2013limited}. $M_2$ is asymptotically $\chi^2$-distributed with $\kappa − \nu$ degrees of freedom ($df$), where $\kappa$ is the reduced number of first- and second-order marginal residuals (i.e. $\kappa = n(n + 1)/2$ for $n$ items) and $\nu$ denotes the number of free parameters. To improve the stability of the $M_2$ statistic, the responses to polytomous items were binarized (i.e. any vs. no maltreatment). Using the $M_2$ statistic, we calculated the root mean square error of approximation (RMSEA), comparative fit index (CFI), and Tucker-Lewis index (TLI) for each model fit. Following convention \citep{hu1999cutoff}, the benchmarks for adequate model fit are RMSEA < 0.08, CFI > 0.90, and TLI > 0.90; in turn, the benchmarks for good model fit are RMSEA < 0.05, CFI > 0.95, and TLI > 0.95.

Given the limitations of SEM-based fit indices for item response models \citep{clark2018model}, we calculated two additional fit indices. First, we used posterior predictive model checking to calculate a $\chi^2_{NC}$ discrepancy measure based on the total score distribution \citep{sinharay2006posterior}. This measure compares the observed and model-predicted proportion of participants at each total score level. Second, to test for local dependence in each pair of items, we calculated Yen's $Q_3$ statistic \citep{yen1984effects} and also the critical $Q_3$-value (i.e. the value above which local dependence is indicated; \citealt{christensen2017critical}). For each model and sample, we simulated 1000 locally independent datasets using the posterior distribution of the model parameters. We then recorded the maximum $Q_3$ value per simulation. The critical $Q_3$-value was defined as the 99th percentile of these max $Q_3$ values.

The goodness-of-fit of the models was compared using Bayesian leave-one out cross- validation (LOO-CV; \citealt{vehtari2017practical}). The LOO-CV measure quantifies the discrepancy between the model and data while taking into account model complexity. LOO-CV values are presented in deviance scale, i.e. smaller values indicate better fit.

@@

In addition to confirmatory factor analysis, we also performed exploratory factor analysis (EFA) on the response data for each sample separately. The purpose for performing EFA after confirmatory factor analysis was to examine whether any of the hypothesized factor structures emerged from the data with fewer \textit{a priori} restrictions \citep{schmitt2018selecting}. For the sake of brevity, the details and results of the EFA are included in the supplement.

@@

The goodness-of-fit indices for each model and sample are summarized in Table \ref{table:cfa_diagnostics}. According to the traditional indices, each model provided at least an acceptable fit to the data. All RMSEA value were <0.05 and most CFI and TLI values were >0.95. The only exception were for the fits of models 1 and 3 to the original sample data where CFI \& TLI > 0.90. In general, model fits were better for the replication sample data than for the original sample. Furthermore, none of the posterior predictive p-values corresponding to the $\chi^2_{NC}$ discrepancy measure exceeded the critical values indicating that all models were able to reproduce the observed distribution of total scores for each sample.

>>>>

% set document class
\documentclass[letterpaper,man,natbib,longtable,floatsintext,12pt]{apa6}

% load packages
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{float}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{booktabs}             % fancy latex tables
\usepackage[export]{adjustbox}    % center wide tables on page
\usepackage{setspace}             % adjust caption linespacing 
\usepackage{multibib}             % references in supplement
\usepackage{hyperref}             % manage hyperlink colors
\usepackage{hanging}

@@

% define title page
\title{Evaluation of the factor structure and structural validity of the Maltreatment and Abuse Chronology of Exposure (MACE) scale}
\shorttitle{Structural validity of the MACE}
\author{Samuel Zorowitz$^{1*}$, Lauri Tuominen$^{2}$}
\affiliation{$^1$Princeton Neuroscience Institute, Princeton University, USA\\$^2$The Royal’s Institute of Mental Health Research, University of Ottawa, Canada\\$^*$Corresponding author: zorowitz@princeton.edu}

@@

{\begin{hangparas}{.25in}{1} \noindent Data availability: All data and analysis materials are publicly available at: \url{https://github.com/szorowi1/mace-irt}. \end{hangparas}}

@@

\noindent \textbf{Purpose}: The Maltreatment and Abuse Chronology of Exposure (MACE) scale is a retrospective self-report measure of childhood maltreatment. Despite increasing use, the factor structure of the MACE has not been previously investigated. As such, the validity of the MACE total and subscale scores are unknown. The objective of the current study was to evaluate the factor structure of the MACE, including the structural validity of its total and subscale scores.

@@

\noindent \textbf{Methods}: We analyzed the response data from two independent samples of participants who completed the MACE (N=1051 \& N=582) using confirmatory factor analysis. We fit a series of models of increasing dimensionality to explore its factor structure. We calculated model-based indices to quantify the reliability of the MACE total and subscale scores, and to test for essential unidimensionality.

@@

\noindent \textbf{Results}: Confirmatory factor analysis indicated that a bifactor S-1 model with two specific factors (peer victimization, reverse-scoring) provided the best and most parsimonious fit to the data. Model-based indices suggested the MACE scale is essentially unidimensional. The MACE total score was determined to be a reliable and valid measure of overall childhood maltreatment. In contrast, MACE subscale scores exhibited little reliable variance after controlling for overall maltreatment.

@@

Childhood maltreatment is a common experience \citep{stoltenborgh2015prevalence} that increases the risk for mental and physical illness  later in life \citep{kessler2010childhood, wegman2009meta} and predicts worse treatment response for multiple psychiatric disorders \citep{nanni2012childhood, thomas2019childhood}. Reliable and valid assessment of childhood maltreatment is an important research goal in order to better understand the link between maltreatment and health outcomes; improve treatment strategies for victims of childhood maltreatment; and design better interventions to prevent maltreatment before it occurs.

There already exist many measures of childhood abuse and maltreatment for researchers to use \citep{saini2019systematic}. One recently developed measure, the Maltreatment and Abuse Chronology of Exposure scale (MACE; \citealt{teicher2015maltreatment}), has a number of notable advantages. First, the MACE measures 10 distinct types of childhood maltreatment, including four types that are seldom included in other scales (e.g., peer victimization, witnessing domestic violence). Second, the items in the MACE were selected using item response theory in order to ensure that each subscale assesses maltreatment experiences of increasing severity. Third, the MACE also measures the timing of maltreatment experiences, which is necessary for investigating how the chronology of abuse shapes development. Finally, the MACE total score (total number of unique maltreatment experiences) exhibits good psychometric properties, including excellent temporal stability, convergent validity with other childhood maltreatment scales, and predictive validity for many psychiatric symptoms \citep{teicher2015maltreatment}. For these reasons, the MACE has been appraised as among the best measures of childhood maltreatment currently available \citep{saini2019systematic, georgieva2022systematic}.

One critical issue for using the MACE in childhood maltreatment research is that its structural validity --- defined as the degree to which an instrument's scores adequately reflect the dimensionality of the construct being measured --- is unknown \citep{saini2019systematic}. This issue is pressing as some researchers have begun to use scores from each of the 10 MACE subscales to study the unique contributions of distinct maltreatment types to developmental outcomes (e.g., \citealt{schalinski2015type, gerke2018childhood, schalinski2019early}). Others have calculated bespoke subscale scores of threat and deprivation exposure using composites of MACE items that respectively measure abuse and neglect experiences (e.g., \citealt{schalinski2018defining, schalinski2019environmental, teicher2018differential}). Because the factor structure of the MACE has not been investigated, the validity of these subscale scores (i.e., the extent to which they reflect particular types of maltreatment as intended) is indeterminate.

Though the MACE measures 10 distinct types of maltreatment, the factor structure that best explains the patterns of responses on the MACE may have fewer than 10 factors given the high co-occurrence of maltreatment experiences. That is, children who have experienced any single type of maltreatment are likely to have experienced multiple other types \citep{dong2004interrelatedness, herrenkohl2009assessing, kessler2010childhood}. As such, it is possible that MACE subscale scores may primarily reflect overall maltreatment rather than any specific type of maltreatment. Bifactor models, a type of hierarchical factor models, are well-suited for resolving this type of concern \citep{bornovalova2020appropriate}. Under a bifactor model, covariation in responses to a set of items are accounted for by a \emph{general} factor, reflecting shared variance across all items, and one or more \emph{specific} factors, reflecting additional shared variance among subsets of items \citep{Reise2012-ql}. Applied to the MACE, bifactor models can aid in determining whether its subscale scores are interpretable as measures of distinct types of maltreatment, or if they instead chiefly reflect general maltreatment.

The main objective of the present study was to investigate the structural validity of the MACE. To do so, we fit a series of confirmatory factor models to the response data from two independent samples of participants who completed the MACE. The first objective of our analyses was to examine the dimensionality of the MACE; that is, to determine whether responses on the MACE are best-explained by a factor structure with 10 factors (one per subscale) or by a more parsimonious factor structure. The second objective was to quantify the reliability of, and sources of variability in, the MACE total and subscale scores in order to determine whether each score adequately measured its construct as intended.

@@

The participants in the current study (N=1633 total) belonged to one of two samples. The first sample of participants (N=1051) are from the original development of the MACE and the details of their recruitment have been reported elsewhere \citep{teicher2015maltreatment}. Briefly, inclusion criteria included being medically healthy, unmedicated, and between 18–25 years of age. This sample will hereafter be referred to as the \textit{original} sample.

The second sample of participants (N=687) were recruited from Prolific Academic (\url{https://www.prolific.co}) to participate in an experiment in August -- October, 2021. Of these, N=582 were retained for analysis (see Exclusion Criteria below). Participants were eligible for participation if they were 18 years or older and resided in the United States or Canada. Participants received monetary compensation for their time (rate: \$12 USD/hr). This study was approved by the Institutional Review Board of the University of Ottawa (\#2021003), and all participants provided informed consent. This sample will hereafter be referred to as the \textit{replication} sample.

The demographics of the samples are summarized in Table \ref{tab:demographics}. The majority of participants in both samples identified as women, but the original sample was composed of proportionally more woman ($z$ = 4.142, $p$ < 0.001). Participants in the original sample were also younger on average ($t$ = -19.313, $p$ < 0.001). The majority of participants identified as White and not as Hispanic or Latino. The proportion of participants identifying as White was not significantly different across samples ($z = 0.417$, $p = 0.677$), nor was the proportion of participants identifying as Hispanic or Latino ($z$ = -1.241, $p$ = 0.214).

\begin{table}[t!]
    \centering
    \begin{tabular*}{\textwidth}{lccc}
    \toprule
    Variable & Original (N=1051) & Replication (N=582) & p-value \\
    \midrule
    \textbf{Gender}, N (\%)                & & & p < 0.001 \\
    \hspace{1em} Women                     & 670 (63.7\%) & 310 (53.3\%) & \\
    \hspace{1em} Men                       & 381 (36.3\%) & 258 (44.3\%) & \\
    \hspace{1em} Transgender or nonbinary         &   0  (0.0\%) &   9  (1.5\%) & \\
    \hspace{1em} Rather not say            &   0  (0.0\%) &   5  (0.9\%) & \\
    \midrule
    \textbf{Age}, yrs                      & & & p < 0.001 \\
    \hspace{1em} Mean (range)              & 23.1 (18 -- 27) & 30.2 (18 -- 80) & \\
    \midrule
    \textbf{Ethnicity}, N (\%)             & & & p = 0.214 \\
    \hspace{1em} Not Hispanic or Latino    & 963 (91.6\%) & 518 (89.0\%) & \\
    \hspace{1em} Hispanic or Latino        &  84  (8.0\%) &  57  (9.8\%) & \\
    \hspace{1em} Rather not say            &   4  (0.4\%) &   7  (1.2\%) & \\
    \midrule
    \textbf{Race}, N (\%)                  & & & p = 0.677 \\
    \hspace{1em} White                     & 806 (76.7\%) & 441 (75.8\%) & \\ 
    \hspace{1em} Asian                     & 103  (9.8\%) &  65 (11.2\%) & \\
    \hspace{1em} Black or African American &  81  (7.7\%) &  31  (5.3\%) & \\
    \hspace{1em} Other                     &  61  (5.8\%) &  28  (4.8\%) & \\
    \hspace{1em} Rather not say            &   0  (0.0\%) &  17  (2.9\%) & \\
    \bottomrule
    \end{tabular*}
    \caption{\normalfont Demographics of the two samples included in this study. The original sample was recruited and reported by \cite{teicher2015maltreatment}. The replication sample was recruited by the current authors for the purposes of this study.}
    \label{tab:demographics}
\end{table}

@@

All participants completed the 52-item version of the MACE \citep{teicher2015maltreatment}. The MACE is a retrospective self-report scale that measures 10 unique types of childhood maltreatment (i.e., parental verbal abuse, physical abuse, nonverbal emotional abuse, sexual abuse, emotional neglect, physical neglect, witnessing violence to siblings, witnessing interparental violence, peer verbal abuse, peer physical abuse). For each item, participants endorsed if they ever experienced a particular event during childhood (e.g., ``Parents or guardians intentionally pushed, grabbed, shoved, slapped, pinched, punched or kicked you'') and, if so, at what ages (up to age 18). Here we only analyze the dichotomous response data (Yes = 1, No = 0). Six of the 52-items --- evenly divided between the emotional and physical neglect subscales --- are reverse-scored (e.g., ``One or more individuals in your family helped you feel important or special''). For these items we followed the scoring procedure set by \cite{teicher2015maltreatment}: a score of 1 is assigned to a response only in the absence of a positive event for all 18 years of childhood. This is in contrast to all remaining items, where a score of 1 is assigned if maltreatment was experienced in at least one year of childhood.

In both the original and replication samples, participants completed a number of additional self-report symptom measures. We did not analyze these data and therefore do not discuss these measures further. For completeness, the additional self-report measures that were collected are reported in the Supplementary Materials.

@@

To ensure data quality in the replication sample, we relied on attention checks embedded in the self-report measures to identify and remove participants engaging in inattentive responding (\citealt{zorowitz2021inattentive}; see the Supplementary Materials for details). The data from N=64 (9.9\%) participants were excluded for failing one or more attention checks, leaving the data from N=582 participants for analysis.

@@

To investigate the factor structure of the MACE, we tested four confirmatory item factor models. The basis of each model is the graded response model for ordered polytomous data \citep{samejima1997graded}. The factor structure for each model is depicted in Figure \ref{fig:models}. The first is a one-factor (undimensional) model where all items load onto a general maltreatment factor (Figure \ref{fig:models}a). The second is a bifactor model where all items load onto a general maltreatment factor and also one of 10 specific factors based on their subscale membership (Figure \ref{fig:models}b). We used this model to evaluate the reliability of the MACE subscale scores.

The third model was motivated by previous research that used the MACE to calculate bespoke subscale scores measuring threat and deprivation experiences (e.g., \citealt{teicher2018differential}). To evaluate those scores, we initially fit a bifactor model where all items loaded onto a general maltreatment factor, and also a threat or neglect specific factor. After observing vanishing loadings on the threat specific factor, however, we instead fit a bifactor S-1 model \citep{eid2017anomalous} where all items loaded onto a general maltreatment factor, and only items measuring neglect load onto a specific neglect factor (Figure \ref{fig:models}c). We used this model to evaluate the reliability of the neglect subscale scores.

@@

The fourth model offers a more parsimonious structure of the MACE where items are organized according to the source of maltreatment: parents (or other adults in the home) and peers. That is, this model represents the hypothesis that two factors are sufficient to summarize responses on the MACE: a primary parental maltreatment factor and a secondary peer victimization factor. Based on preliminary model fits, we suspected the presence of a tertiary methods factor affecting the six reverse-scored items. We initially fit a bifactor model in which all items loaded onto a general maltreatment factor and one of three specific factors (i.e., parental maltreatment, peer victimization, reverse-scoring). After observing vanishing loadings on the first specific factor, however, we instead fit a bifactor S-1 model where all items load onto a general maltreatment factor, and the peer victimization and reverse-scored items additionally load onto their own specific factors (Figure \ref{fig:models}d). We fit this model to test a more parsimonious factor structure of the MACE, and to evaluate the reliability of peer victimization subscale scores.

We fit each model separately to the data from the original and replication samples. We did so because, prior to analysis, we detected uniform differential item functioning (DIF) by sample for 11 items (see the Supplementary Materials for details). Also prior to analysis, we identified a total of 12 pairs and triplets of MACE items with response dependence. For example, a person endorsing item 8 (``Parents or guardians hit you so hard that you received medical attention'') is all but certain to endorse item 7 (``Parents or guardians hit you so hard that it left marks for more than a few minutes''). To prevent bias in the estimation of factor loadings due to response dependence \citep{reise2013applying}, we followed best-practice recommendations and combined each set of dependent items into a single polytomous item \citep{marais2008formalizing}. Thus, we reduced the MACE from 52 to 39 independent items (Table \ref{tab:dependence}).

@@

We relied on multiple indices to judge the fit of the models to the data. First we calculated three traditional goodness-of-fit statistics based on the ordinal $M_2$ statistic \citep{cai2013limited}. Specifically, we calculated the root mean square error of approximation (RMSEA), comparative fit index (CFI), and Tucker-Lewis index (TLI) for each model fit. Following convention \citep{hu1999cutoff}, we defined the benchmarks for adequate model fit as RMSEA < 0.08, CFI > 0.90, and TLI > 0.90, and the benchmarks for good model fit as RMSEA < 0.05, CFI > 0.95, and TLI > 0.95.

Given the limitations of SEM-based fit indices for item response models (e.g., \citealt{clark2018model}), we calculated two additional fit indices. First, we used posterior predictive model checking to calculate a $\chi^2_{NC}$ discrepancy measure based on the total score distribution \citep{sinharay2006posterior}. This measure compares the observed and model-predicted proportion of participants at each total score level. Second, to test for local dependence in each pair of items, we calculated Yen's $Q_3$ statistic \citep{yen1984effects} and also the critical $Q_3$-value (i.e., the value above which local dependence is indicated; \citealt{christensen2017critical}). Briefly, for each model and sample, we simulated 1000 locally-independent datasets using the posterior distribution of model parameters. We then recorded the maximum $Q_3$ value per simulation. The critical $Q_3$-value was defined as the 99th percentile of these max $Q_3$ values.

The goodness-of-fit of the models was compared using Bayesian leave-one-out cross-validation (LOO-CV; \citealt{vehtari2017practical}). LOO-CV quantifies the discrepancy between the model and data while taking into account model complexity. LOO-CV values are reported here in deviance scale (i.e., smaller values indicate better fit).

@@

In addition to the confirmatory factor models, we also performed exploratory factor analysis (EFA) on each sample's data separately. The motivation for performing EFA was to examine whether any of the confirmatory factor structures emerged from the data with fewer \textit{a priori} restrictions \citep{schmitt2018selecting}. The details and results of the EFA are included in the Supplementary Materials.

@@

The goodness-of-fit indices for each sample and model are summarized in Table \ref{table:cfa_diagnostics}. According to the traditional indices, each model provided at least an acceptable fit to the data. All RMSEA value were <0.05 and most CFI and TLI values were >0.95. The only exception were for the fits of models 1 and 3 to the original sample data where CFI and TLI > 0.90. Furthermore, none of the posterior predictive $p$-values from the $\chi^2_{NC}$ discrepancy measure exceeded the critical values, indicating that all models were able to reproduce the observed total score distribution for each sample.
